---
title: "CSF p-tau at baseline ~ TC pgs in ADNI"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 3
    fig_width: 6.5
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input
Relative to project base directory:
```{r}
pheno.file <- "data/adni/adni_summary_data.txt"
```

## Libraries
```{r}
library(tidyverse)
library(gridExtra)

source(paste(path.to.project, "scripts/project_functions.R", sep = ""))
```

## Data
```{r}
pheno <- read.within.project.h(pheno.file)
summary(pheno)
```

### Filter data for presence of CSF data & family history
```{r}
csf <- filter(pheno, !is.na(bl.csf.ptau) & fh.ad == 1 & !is.na(tc.pgs)) %>%
    mutate(csf.ptau.tau = bl.csf.ptau/bl.csf.tau) %>%
    select(iid, tc.pgs, sex = sex.f, apoe4.status, age = bl.age, statin = statin.sc.bl, 
           bl.dx, csf.ptau.tau, tc.mM = sc.tc.mM, PC1:PC10)
summary(csf)
```

Missing data for 31 individuals on statin use. These are excluded because included as a variable in the regression. 

### Factorize data
```{r}
csf.f <- filter(csf, !is.na(statin))
csf.f$apoe4.status <- factor(csf.f$apoe4.status, levels = c(0, 1), labels = c("APOE4_neg", "APOE4_pos"))
csf.f$sex <- factor(csf.f$sex, levels = c(0, 1), labels = c("Male", "Female"))
csf.f$statin <- factor(csf.f$statin, levels = c(0, 1), labels = c("Statin_neg", "Statin_pos"))

# data for descriptives
m.dt <- filter(csf, !is.na(statin))
m.dt$sex <- factor(m.dt$sex, levels = c(0, 1), labels = c("Male", "Female"))
```


## Descriptives
```{r}
desc.stat <- sum.stat(m.dt, strat_by = "sex", 
         cont_variables = c("age", "tc.mM", "tc.pgs", "csf.ptau.tau"), 
         disc_variables = c("statin", "apoe4.status"))
desc.stat[, " "] <- sapply(desc.stat$p, function(x) if(x <= 0.05) {"*"} else {""})
desc.stat
```

## EDA

#### TC-PGS distribution
```{r}
grid.arrange(
    ggplot(csf.f, aes(tc.pgs)) + 
        geom_histogram(data = filter(csf.f, sex == "Female"), bins = 10, fill = "#f5dd4dff") +
        geom_histogram(data = filter(csf.f, sex == "Male"), bins = 10, fill = "#1f3c6dff") +
        my.theme, 
    
    ggplot(csf.f, aes(tc.pgs, fill = sex)) +
        geom_density(alpha = 0.5) +
        scale_fill_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme,
    
    ncol = 2
)
```

#### CSF distribution
```{r}
grid.arrange(
    ggplot(csf.f, aes(csf.ptau.tau)) + 
        geom_histogram(data = filter(csf.f, sex == "Female"), bins = 10, fill = "#f5dd4dff") +
        geom_histogram(data = filter(csf.f, sex == "Male"), bins = 10, fill = "#1f3c6dff") +
        my.theme, 
    
    ggplot(csf.f, aes(csf.ptau.tau, fill = sex)) +
        geom_density(alpha = 0.5) +
        scale_fill_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme,
    
    ncol = 2
)
```
Data look normally distributed. 

## TC-PGS effect on PET data

### Linear regression - All individuals

**Statistics**
```{r}
formula.1 <- as.formula("csf.ptau.tau ~ PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10+age + apoe4.status + statin * sex * tc.pgs")
linreg.1 <- lm(formula.1, data = csf.f)
```

Assumptions:
```{r, fig.height=6}
par(mfrow = c(2, 2))
plot(linreg.1)
```

Results:
```{r}
summary(linreg.1)
```

**Plotting**
```{r, echo=FALSE}
ggplot(csf.f, aes_string(x = "tc.pgs", y = "csf.ptau.tau", color = "sex")) +
    geom_point(size = 3) +
    geom_smooth(method = "lm", se = FALSE) +
    scale_color_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
    facet_grid(~statin) +
    my.theme
```

## Save data
```{r}
save(list = c("csf.f", "desc.stat"), file = "adni_csf_ptau.tau.RData")
```

