---
title: "Log of 'adni_csf_data.txt'"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

### Raw data source
Raw data was downloaded from http://adni.loni.usc.edu/. Observe that while most other data was downloaded 2015-12-03, CSF data were downloaded 2018-06-22. 
```{r, echo=FALSE}
raw.dir <- "~/Research/Cohort_data/ADNI/Phenotypic_data/Raw_data/" 
```
Collected from: `r raw.dir`

### Raw data files
```{r input}
# Raw data from outsided project
key.file <- "2015-12-03/ROSTER.csv"
csf.file <- "2018-06-22/UPENNBIOMK9_04_19_17.csv"
```

## Libraries
```{r}
library(tidyverse)
```

## Key data
```{r}
key <- read.csv(paste(raw.dir, key.file, sep = ""), header = TRUE) %>%
    select(RID, IID = PTID)
head(key)
```

## CSF data

### Read & clean up of data
CSF data column names:
```{r}
colnames(read.csv(paste(raw.dir, csf.file, sep = ""), header = TRUE))
csf <- read.csv(paste(raw.dir, csf.file, sep = ""), header = TRUE) %>%
    merge.data.frame(key) %>%
    select(IID, VISCODE2, EXAMDATE, ABETA, TAU, PTAU, COMMENT) %>%
    unique()
```

Overview of data:
```{r}
summary(csf)
```

Replace empty cells with NAs
```{r}
csf[csf == ""] <- NA
summary(csf)
```

Clean up csf data:
```{r}
unique(csf$COMMENT)

csf.v2 <- csf %>%
    separate(COMMENT, c(paste("V", 1:6, sep = "")), sep = " ") %>%
    mutate(csf.tau = as.numeric(as.character(TAU)), 
           csf.ptau = as.numeric(as.character(PTAU)), 
           abeta.n = as.numeric(as.character(ABETA)), 
           abeta.2 = as.numeric(V5)) 

csf.v2$csf.ab42 <- apply(select(csf.v2, abeta.2, abeta.n), 1, function(x) 
    if(!is.na(x[1])) {x[1]} else {x[2]})
```

Unique labels in "VISCODE2":
```{r}
unique(csf.v2$VISCODE2)
```

To create a numeric time variable, "nv" in "VISCODE2" are replaced with NAs and "bl" with 0 - this will be the **"csf.time"**, variable.
```{r}
cleaned.data <- csf.v2 %>%
    mutate(VISCODE.rm1 = gsub("bl", 0, VISCODE2), 
           VISCODE.rm2 = gsub("nv", NA, VISCODE.rm1),
           csf.time = as.numeric(gsub("m", "", VISCODE.rm2))) %>% 
    select(IID, VISCODE2, csf.time, EXAMDATE, contains("csf"))
head(cleaned.data)
```

### Baseline data 
```{r}
csf.bl <- filter(cleaned.data, csf.time == 0) %>%
    select(IID, bl.csf.tau = csf.tau, bl.csf.ptau = csf.ptau, bl.csf.ab42 = csf.ab42) %>%
    unique()

# check for duplicates
sum(duplicated(csf.bl$IID))
```

### Last visit data
```{r}
# Last csf visit ####
csf.lv <- filter(cleaned.data, !is.na(csf.time)) %>% 
    group_by(IID) %>%
    summarise(lv.csf.time = max(csf.time, na.rm = TRUE)) %>%
    unique()

# Last csf status ####
csf.lvstatus <- merge.data.frame(cleaned.data, csf.lv) %>%
    filter(csf.time == lv.csf.time) %>%
    select(IID, lv.csf.tau = csf.tau, lv.csf.ptau = csf.ptau, lv.csf.ab42 = csf.ab42, lv = VISCODE2, lv.csf.time) %>%
    unique()

# Check for duplicates
duplicated(csf.lvstatus$IID) %>% sum()
```

```{r}
csf.final <- cleaned.data %>%
    select(IID, VISCODE2, EXAMDATE, csf.time, csf.tau, csf.ptau, csf.ab42) %>%
    merge.data.frame(csf.bl) %>%
    merge.data.frame(csf.lvstatus)
summary(csf.final)
```

## Write data
```{r}
write.table(csf.final, file = "adni_csf_data.txt", 
            row.names = FALSE, quote = FALSE, sep = "\t")
```