---
title: "Log of 'adni_statin_data.txt'"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

### Raw data source
Raw data was downloaded from http://adni.loni.usc.edu/ 2015-12-03 unless otherwise stated.
```{r, echo=FALSE}
raw.dir <- "~/Research/Cohort_data/ADNI/Phenotypic_data/Raw_data/2015-12-03/" 
```
Collected from: `r raw.dir`

### Raw data files
```{r input}
key.file <- "ROSTER.csv"
reg.file <- "REGISTRY.csv"
med.file <- "RECCMEDS.csv"
```

List of statin brand names:
```{r}
statin.brand.name.file <- "~/Research/Other_data/statin_brand_names.xlsx"
```

## Libraries
```{r}
library(tidyverse)
library(openxlsx)
```

## Key data
```{r}
key <- read.csv(paste(raw.dir, key.file, sep = ""), header = TRUE) %>%
    select(RID, IID = PTID)
head(key)
```

## Registry data
```{r}
colnames(read.csv(paste(raw.dir, reg.file, sep = ""), header = TRUE))
reg <- read.csv(paste(raw.dir, reg.file, sep = ""), header = TRUE) %>%
    merge.data.frame(key) %>%
    select(Phase, RID, IID, VISCODE2, EXAMDATE) %>%
    unique()
summary(reg)

# Replace emtpy cells with NAs
reg[reg == ""] <- NA

filter(reg, is.na(VISCODE2))

reg.cc <- reg[complete.cases(reg), ] %>%
    mutate(reg.examdate = as.Date(EXAMDATE)) %>%
    select(-EXAMDATE)
summary(reg.cc)


```

## Medication data

#### Brand names

```{r}
brand <- read.xlsx(statin.brand.name.file, colNames = FALSE) 
brand$X2
brand.name <- sapply(brand$X2, function(x) strsplit(x, split = ", ")) %>% 
    unlist(recursive = TRUE, use.names = FALSE)
brand.name[order(brand.name)]
```

#### ADNI medication data
```{r}
colnames(read.csv(paste(raw.dir, med.file, sep = ""), header = TRUE))
med <- read.csv(paste(raw.dir, med.file, sep = ""), header = TRUE) %>%
    select(Phase, RID, VISCODE2, CMMED, CMBGN, CMEND) %>%
    unique()
summary(med)

# Replace empty spaces and -4 with NAs
med[med == "-4" | med == -4 | med == ""] <- NA
summary(med)

med.cl <- merge.data.frame(reg.cc, med) %>%
    select(PHASE = Phase, RID, IID, drug.date = reg.examdate, VISCODE2, CMMED, CMBGN, CMEND) %>%
    unique() %>%
    mutate(MEDICATION = tolower(CMMED))
head(med.cl)
```

#### Extracting brand names from ADNI data
```{r}
unique(med.cl$CMMED)

statins <- med.cl$MEDICATION[grepl("statin", med.cl$MEDICATION)] %>%
    unique()
statins
statins.brand <- med.cl$MEDICATION[grepl(paste(brand.name, collapse = "|"), med.cl$MEDICATION)] %>%
    unique()

statins.cl <- c(statins, "provachol", statins.brand)

statins.df <- filter(med.cl, MEDICATION %in% c(NA, statins.cl)) %>%
    select(PHASE, IID, VISCODE2, drug.date, MEDICATION, CMBGN, CMEND) %>%
    mutate(statin = as.numeric(!is.na(MEDICATION))) %>%
    unique()
head(statins.df)
```

#### Screening visit statin data
```{r}
sc <- filter(statins.df, VISCODE2 == "sc") %>%
    select(IID, statin.sc.date = drug.date, statin) %>%
    unique()

sum(duplicated(sc$IID))
sc.dup <- filter(sc, IID %in% sc$IID[duplicated(sc$IID)])

sc.unique <- group_by(sc, IID, statin.sc.date) %>% # getting rid of duplicate values
    summarise(statin.sc = sum(statin)) %>%
    unique()

unique(sc.unique$statin.sc)
sum(duplicated(sc.unique$IID))
```

#### Baseline visit statin data
```{r}
bl <- filter(statins.df, VISCODE2 == "bl") %>%
    select(IID, statin.bl.date = drug.date, statin) %>%
    unique()

sum(duplicated(bl$IID))
filter(bl, IID == "036_S_4389")

bl.unique <- group_by(bl, IID, statin.bl.date) %>% # getting rid of duplicate values
    summarise(statin.bl = sum(statin)) %>%
    unique()

unique(bl.unique$statin.bl)
sum(duplicated(bl.unique$IID))
```

#### Statin ever use
```{r}
ev <- group_by(statins.df, IID) %>%
    summarize(temp = sum(statin, na.rm = TRUE)) %>%
    mutate(statin.everuse = as.numeric(temp > 0))
sum(duplicated(ev$IID))
```

## Merge data
```{r}
m <- merge.data.frame(statins.df, sc.unique, all.x = TRUE) %>%
    merge.data.frame(bl.unique, all.x = TRUE) %>%
    merge.data.frame(ev, all.x = TRUE)
```

## Write data
```{r}
write.table(m, file = "adni_statin_data.txt", row.names = FALSE, quote = FALSE, sep = "\t")
```
