---
title: "Log of 'adni_demog_data'"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

### Raw data source
Raw data was downloaded from http://adni.loni.usc.edu/ 2015-12-03 unless otherwise stated.
```{r, echo=FALSE}
raw.dir <- "~/Research/Cohort_data/ADNI/Phenotypic_data/Raw_data/2015-12-03/" 
```
`r raw.dir`

### Raw data files
```{r input}
key.file <- "ROSTER.csv"
fhq.file <- "FHQ.csv" # parental AD history
sib.file <- "RECFHQ.csv" # sibling AD history
```

## Libraries
```{r}
library(tidyverse)
```

## Data

#### Key data
```{r}
key <- read.csv(paste(raw.dir, key.file, sep = ""), header = TRUE) %>%
    select(RID, IID = PTID)
```

#### Parental data

Parental history of AD is located in columns "FHQDADAD" & "FHQMOMAD" (for dad and mom respectively), with the following values: 1=Yes; 0=No; 2=Don't Know.
```{r}
par <- read.csv(paste(raw.dir, fhq.file, sep = ""), header = TRUE) %>%
    merge.data.frame(key) %>%
    select(IID, dad.ad = FHQDADAD, mom.ad = FHQMOMAD) %>%
    unique()
summary(par)
```

#### Sibling data

AD history in siblings are recorded in "FHQSIBAD" column with the following values: 1=Yes; 0=No; 2=Don't Know

```{r}
sib <- read.csv(paste(raw.dir, sib.file, sep = ""), header = TRUE) %>%
    merge.data.frame(key) %>%
    select(IID, sib.ad = FHQSIBAD) %>%
    unique()
summary(sib)
```

#### Merged data
```{r}
m <- merge.data.frame(par, sib, all = TRUE) %>%
    unique()

apply(m[, -1], 2, unique)
```

In addition to 0, 1 and 2, data also contains "-1" and "-4". These will be treated as missing similar to 2. 

#### Family history variable
Creating fh.ad column coded 1 for known AD family history in either mom, dad or sibling. Others are coded NA.
```{r}
m$fh.ad <- apply(m[, -1], 1, function(x) {
    if("1" %in% x) {1}
    else {NA}
})

m.f <- m %>%
    filter(fh.ad == 1) %>%
    select(IID, fh.ad) %>%
    unique()
```

#### Duplicate check
```{r}
sum(duplicated(m.f$IID))
```

## Write data
```{r}
write.table(m.f, file = "adni_fh_data.txt", 
            row.names = FALSE, quote = FALSE, sep = "\t")
```