---
title: 'ADNI: Disease status data'
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

### Raw data source
Raw data was downloaded from http://adni.loni.usc.edu/ 2015-12-03 unless otherwise stated.
```{r, echo=FALSE}
raw.dir <- "~/Research/Cohort_data/ADNI/Phenotypic_data/Raw_data/2015-12-03/" 
```
Collected from: `r raw.dir`

### Raw data files
```{r, echo=FALSE}
key.file <- "ROSTER.csv"
raw.pheno.file <- "DXSUM_PDXCONV_ADNIALL.csv"
```
* `r raw.pheno.file`
* `r key.file`

## Libraries
```{r}
library(tidyverse)
```

## Inspect raw data
Columns in `r raw.pheno.file`:
```{r, echo=FALSE}
colnames(read.csv(paste(raw.dir, raw.pheno.file, sep = "")))
```

Columns in `r key.file`:
```{r, echo=FALSE}
colnames(read.csv(paste(raw.dir, key.file, sep = "")))
```

## Read, merge & select data:

#### Raw disease data
```{r}
key <- read.csv(paste(raw.dir, key.file, sep = "")) %>%
    select(RID, IID = PTID)
head(key)

rawdata <- read.csv(paste(raw.dir, raw.pheno.file, sep = "")) %>%
    merge.data.frame(key) %>%
    select(RID, IID, VISCODE2, EXAMDATE, DXCHANGE, DXCURREN) %>%
    unique() 
head(rawdata)
```

Unique labels in "VISCODE2":
```{r}
unique(rawdata$VISCODE2)
```

#### Clean up of data
To create a numeric time variable, "sc" and "uns1" in "VISCODE2" are replaced with NAs and "bl" with 0 - this will be the **"dx.time"**, variable.
```{r}
cleaned.data <- rawdata %>%
    mutate(VISCODE.rm1 = gsub("bl", 0, VISCODE2), 
           VISCODE.rm2 = gsub("sc", NA, VISCODE.rm1),
           VISCODE.rm3 = gsub("uns1", NA, VISCODE.rm2),
           dx.time = as.numeric(gsub("m", "", VISCODE.rm3))) %>% 
    select(RID, IID, VISCODE2, dx.time, EXAMDATE, DXCHANGE, DXCURREN)
head(cleaned.data)
```

**No of unique IDs:**
```{r, echo=FALSE}
length(unique(cleaned.data$IID))
```

**dx variable**
Currently disease information is located in two columns: "DXCHANGE" & "DXCURREN" and neither seem to be complete. Thus, data from both columns will be combined into a new variable **dx** and converted to a format where 1 = Control, 2 = MCI, and 3 = AD.
```{r}
# Combine columns
cleaned.data$dx <- apply(cleaned.data[, c("DXCHANGE", "DXCURREN")], 1, function(x) 
    if(!is.na(x[2])) {x[2]}
    else if(is.na(x[2] & !is.na(x[1]))) {x[1]}
    else {NA})

# Replace conversion # with status # ####
cleaned.data$dx[cleaned.data$dx == 7 | cleaned.data$dx == 9] <- 1
cleaned.data$dx[cleaned.data$dx == 4 | cleaned.data$dx == 8] <- 2
cleaned.data$dx[cleaned.data$dx == 5 | cleaned.data$dx == 6] <- 3
```

#### Baseline data 
```{r}
dx.bl <- filter(cleaned.data, dx.time == 0) %>%
    select(IID, bl.dx = dx) %>%
    unique()

# check for duplicates
sum(duplicated(dx.bl$IID))
```

#### Last visit data
```{r}
# Last dx visit ####
dx.lv <- filter(cleaned.data, !is.na(dx)) %>% 
    group_by(IID) %>%
    summarise(lv.dx.time = max(dx.time, na.rm = TRUE)) %>%
    unique()

# Last dx status ####
dx.lvstatus <- merge.data.frame(cleaned.data, dx.lv) %>%
    filter(dx.time == lv.dx.time) %>%
    select(IID, lv.dx = dx, lv = VISCODE2, lv.dx.time) %>%
    unique()

# Check for duplicates
duplicated(dx.lvstatus$IID) %>% sum()
```

#### Status change
** Function to determine if status have changed over time **
```{r}
f_reversion <- function(y) {
    noNA <- as.integer(y[which(!is.na(y))])
    conv.rev <- diff(noNA)
    rev <- sum(conv.rev < 0)
    conv <- sum(conv.rev > 0)
    if (conv == 0 & rev == 0) {
        "No"
    } else if (conv != 0 & rev == 0) {
        "Conv"
    } else if (conv == 0 & rev != 0) {
        "Rev"
    } else if (conv != 0 & rev != 0) {
        "Conv/Rev"
    }
}
```

** Adding column indicating whether reversion/conversion has occurred **
```{r}
dx.w <- cleaned.data %>%
    filter(!is.na(dx.time)) %>%
    mutate(KEY = paste("m", dx.time, sep = "_")) %>%
    select(IID, KEY, dx) %>%
    unique() %>%
    spread(KEY, value = dx) %>%
    select(IID, m_0, m_6, m_12, m_18, m_24, m_36, m_48, m_60, m_72, m_84, m_96, m_108, m_120)

dx.w$dx.change <- apply(dx.w[, 2:ncol(dx.w)], 1, f_reversion)
```

#### Conversion data for Kaplan-Meier analyses
```{r}
# function to find visit where change was recorded
conv.visit.f <- function(x, conversion.value) {
    if(sum(x %in% conversion.value, na.rm = TRUE) == 0) {names(x)[max(which(!is.na(x)))]}
    else { names(x)[min(which(x %in% conversion.value))]}
}
```

**CI conversion**
Conversion to MCI/AD
```{r}
dx.w$ci.conv.time <- apply(select(dx.w, m_0:m_120), 1, function(y) {
    gsub("m_", "", conv.visit.f(y, c(2, 3))) %>%
        as.character() %>%
        as.numeric()
    })
dx.w$ci.conv <- apply(select(dx.w, m_0:m_120), 1, function(z) {
    as.numeric(sum(z %in% c(2, 3), na.rm = TRUE) > 0) 
})
```

**AD conversion**
Conversion to MCI/AD
```{r}
dx.w$ad.conv.time <- apply(select(dx.w, m_0:m_120), 1, function(y) {
    gsub("m_", "", conv.visit.f(y, 3)) %>%
        as.character() %>%
        as.numeric()
    })
dx.w$ad.conv <- apply(select(dx.w, m_0:m_120), 1, function(z) {
    as.numeric(sum(z %in% 3, na.rm = TRUE) > 0) 
})
```

#### Merge data
```{r}
dx.m <- select(cleaned.data, IID, VISCODE2, dx.time, EXAMDATE, dx) %>%
    merge.data.frame(dx.bl) %>%
    merge.data.frame(dx.lvstatus) %>%
    merge.data.frame(select(dx.w, IID, dx.change, contains("conv")))
head(dx.m)
```

#### Binary status, reversion cases excluded
```{r}
dx.m$ad.bin <- apply(dx.m[, c("dx", "dx.change")], 1, function(x) 
    if(grepl("Rev", x[2]) | is.na(x[1])) {NA}
    else {as.numeric(x[1]) - 1})
dx.m$ad.bin[dx.m$ad.bin == 1] <- NA
dx.m$ad.bin[dx.m$ad.bin == 2] <- 1

dx.m$ci.bin <- apply(dx.m[, c("dx", "dx.change")], 1, function(x) 
    if(grepl("Rev", x[2]) | is.na(x[1])) {NA}
    else {as.numeric(x[1]) - 1})
dx.m$ci.bin[dx.m$ci.bin == 2] <- 1

summary(dx.m)
```

## Write data
```{r}
write.table(dx.m, file = "adni_dx_data.txt", row.names = FALSE, quote = FALSE, sep = "\t")
```

