---
title: "Log of 'adni_demog_data'"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

### Raw data source
Raw data was downloaded from http://adni.loni.usc.edu/ 2015-12-03 unless otherwise stated.
```{r, echo=FALSE}
raw.dir <- "~/Research/Cohort_data/ADNI/Phenotypic_data/Raw_data/2015-12-03/" 
```
`r raw.dir`

### Raw data files
```{r input}
key.file <- "ROSTER.csv"
demog.file <- "PTDEMOG.csv"
apoe.file <- "APOERES.csv"
```

## Libraries
```{r}
library(tidyverse)
```

```{r}
key <- read.csv(paste(raw.dir, key.file, sep = ""), header = TRUE) %>%
    select(RID, IID = PTID)
head(key)

demog <- read.csv(paste(raw.dir, demog.file, sep = ""), header = TRUE) %>%
    merge.data.frame(key)
head(demog)

apoe <- read.csv(paste(raw.dir, apoe.file, sep = ""), header = TRUE) %>%
    merge.data.frame(key)
head(apoe)
```

## Demographic data

#### Sex
```{r}
demog[demog == -5 | demog == -4] <- NA

sex <- demog %>%
    mutate(SEX_F = PTGENDER - 1) %>%
    select(IID, SEX_F) %>%
    filter(!is.na(SEX_F)) %>%
    unique()
nrow(sex)
```

#### Date of birth

Raw data only contains year and month, thus, will put day as the 1st for all indiviuals.
```{r}
dob <- demog %>%
    mutate(dob = as.Date(paste(PTDOBYY, PTDOBMM, "01", sep = "-"))) %>%
    filter(!is.na(dob)) %>%
    select(IID, dob) %>%
    unique()
nrow(dob)
summary(dob)
```

#### Education
```{r}
edu <- demog %>%
    mutate(EDUCATION = PTEDUCAT) %>%
    select(IID, EDUCATION) %>%
    filter(!is.na(EDUCATION)) %>%
    unique()
nrow(edu)
```

## APOE data
```{r}
apoe.cl <- apoe %>%
    mutate(apoe.geno = paste(APGEN1, APGEN2, sep = ":"), 
           apoe4.status = as.numeric(grepl("4", apoe.geno))) %>%
    select(IID, apoe.geno, apoe4.status) %>%
    filter(!is.na(apoe.geno)) %>%
    unique()
head(apoe.cl)
```

## Merge data
```{r}
m <- Reduce(function(x, y) merge(x, y, all = TRUE), list(sex, dob, edu, apoe.cl)) %>%
    filter(!is.na(IID)) %>%
    unique()
head(m)
```

#### Check duplicate IIDs
```{r}
length(unique(m$IID)) == nrow(m)

dup.iids <- m$IID[duplicated(m$IID)]
filter(m, IID %in% dup.iids)
```

Duplicated IIDs are due to differing years of education. Replace with the mean of the two values:
```{r}
m.dup <- filter(m, IID %in% dup.iids) %>%
    group_by(IID) %>%
    summarise(mean.edu = mean(EDUCATION))

m[m$IID == m.dup$IID[1], "EDUCATION"] <- m.dup$mean.edu[1]
m[m$IID == m.dup$IID[2], "EDUCATION"] <- m.dup$mean.edu[2]
m[m$IID == m.dup$IID[3], "EDUCATION"] <- m.dup$mean.edu[3]
m[m$IID == m.dup$IID[4], "EDUCATION"] <- m.dup$mean.edu[4]
m[m$IID == m.dup$IID[5], "EDUCATION"] <- m.dup$mean.edu[5]

filter(m, IID %in% dup.iids)

m.unique <- unique(m)
summary(m.unique)
```

## Write data
```{r}
write.table(m.unique, file = "adni_demog_data.txt", 
            row.names = FALSE, quote = FALSE, sep = "\t")
```