---
title: "Log of 'adni_summary_data.txt'"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input
Cleaned data: 
```{r input}
demog.file <- "extracted_from_rawdata/adni_demog_data.txt"
statin.file <- "extracted_from_rawdata/adni_statin_data.txt"
tc.file <- "extracted_from_rawdata/adni_tc_data.txt"
csf.file <- "extracted_from_rawdata/adni_csf_data.txt"
dx.file <- "extracted_from_rawdata/adni_dx_data.txt"
fh.file <- "extracted_from_rawdata/adni_fh_data.txt"
smoke.file <- "extracted_from_rawdata/adni_smoking_data.txt"
```

TC-PGS data (p-value cut-off 1e-6):
```{r}
pgs.file <- "analyses/adni/v2_prad_snps/1_pg_scoring/score_files/adni_tc_pgs.S09.profile"
```

Genetic PC data from outside project:
```{r}
pca.file <- "~/Research/Cohort_data/ADNI/Genetic_data/ADNI/ADNI_EUR_PCAv3/ADNI_EUR_v3_PCs.txt"
```

## Libraries
```{r}
library(tidyverse)
```

## Data
The PC file contains the individuals with genetic data that can be scored (PGS) and are thus the starting point for this cohort. 

#### Genetic PC data
```{r}
pca <- read.table(pca.file, header = TRUE)
head(pca)
summary(pca)
```

#### TC-PGS data
```{r}
pgs <- read.table(paste(path.to.project, pgs.file, sep = ""), header = TRUE)
summary(pgs)


```
Due to missingness of data, this score contains betweem `r min(pgs$CNT)/2` and `r max(pgs$CNT)/2` SNPs.

**Standardize score**
```{r}
pgs$tc.pgs <- scale(pgs$SCORE)

pgs.cl <- select(pgs, IID, tc.pgs)
```

#### Demographic data
```{r}
demog <- read.table(demog.file, header = TRUE)
summary(demog)
sum(duplicated(demog$IID))
```

#### Statin data
```{r}
statin <- read.table(statin.file, header = TRUE, sep = "\t")
summary(statin)

statin.bl <- select(statin, IID, statin.sc.date, statin.sc, statin.bl.date, statin.bl, statin.everuse) %>%
    unique()
```

#### TC data
```{r}
tc <- read.table(tc.file, header = TRUE, sep = "\t")
summary(tc)

tc.sc <- filter(tc, VISCODE2 == "sc") %>%
    select(IID, tc.date, tc.mM) %>%
    unique()
sum(duplicated(tc.sc$IID))
```

There are multiple values for individuals at the same visit. These will be averaged.
```{r}
tc.sc.g <- group_by(tc.sc, IID) %>%
    summarise(sc.tc.mM = mean(tc.mM)) %>%
    merge.data.frame(select(tc.sc, IID, tc.date)) %>%
    select(IID, tc.date, sc.tc.mM) %>%
    unique()
sum(duplicated(tc.sc.g$IID))
```

#### CSF data
```{r}
csf <- read.table(csf.file, header = TRUE)
head(csf)

csf.age <- merge.data.frame(csf, select(demog, IID, dob)) %>%
    mutate(age = as.numeric(as.Date(EXAMDATE) - as.Date(dob))/365.25)

csf.bl <- filter(csf.age, VISCODE2 == "bl") %>%
    select(IID, csf.bl.age = age, contains("bl"))
csf.lv <- filter(csf.age, csf.time == lv.csf.time) %>%
    select(IID, lv.csf.age = age, lv.csf = lv, lv.csf.tau, lv.csf.ptau, lv.csf.ab42)
```

#### Disease data
```{r}
dx <- read.table(dx.file, header = TRUE)
head(dx)

dx.f <- dx %>% 
    merge.data.frame(select(demog, IID, dob)) %>%
    mutate(age = as.numeric(as.Date(EXAMDATE) - as.Date(dob))/365.25)

dx.bl <- filter(dx.f, VISCODE2 == "bl") %>%
    select(IID, bl.age = age, bl.dx, bl.ad = ad.bin, bl.ci = ci.bin)
dx.lv <- filter(dx.f, lv.dx.time == dx.time) %>%
    select(IID, lv.age = age, lv.dx.time, lv.dx, lv.ad = ad.bin, lv.ci = ci.bin) %>%
    unique()
dx.conv <- select(dx, IID, contains("conv")) %>%
    unique()
```

#### AD family history data
```{r}
fh <- read.table(fh.file, header = TRUE)
head(fh)
```

#### Smoking data
```{r}
smoke <- read.table(smoke.file, header = TRUE)
head(smoke)
```

#### Merge data
```{r}
my.merge <- function(x,y) {
    merge.data.frame(x, y, by = "IID", all.x = TRUE)
}

m <- Reduce(my.merge, list(pca, pgs.cl, demog, fh, statin.bl, smoke, tc.sc.g, dx.bl, dx.lv, dx.conv, csf.bl, csf.lv)) %>%
    unique()
```

**Duplicated values**
```{r}
sum(duplicated(m$IID))
summary(m)
```

```{r}
filter(m, IID == "031_S_0830")
```

Individual 031_S_0830 have two values for last visit age (lv.age) but they are fairly similar (75.90 vs 75.89). Thus, the mean of these two will be used. 
```{r}
m.g <- group_by(m, IID) %>%
    summarise(lv.age = mean(lv.age)) %>%
    merge.data.frame(select(m, - lv.age)) %>%
    unique()
filter(m.g, IID == "031_S_0830") %>% select(IID, lv.age)
```

## New variables

#### ATN status
ATN status is determined based on Hansson et al., 2018 (Alz & Dem) with a amyloid positive status defined as a p-tau/AB42 ratio >= 0.028

```{r}
m.g$bl.ptau.ab42 <- m.g$bl.csf.ptau/m.g$bl.csf.ab42
m.g$bl.atn.a <- as.numeric(m.g$bl.ptau.ab42 >= 0.028)

m.g$lv.ptau.ab42 <- m.g$lv.csf.ptau/m.g$lv.csf.ab42
m.g$lv.atn.a <- as.numeric(m.g$lv.ptau.ab42 >= 0.028)
```

#### Age at tc and statin
```{r}
m.g$tc.age <- as.numeric(as.Date(m.g$tc.date) - as.Date(m.g$dob))/365.25
m.g$statin.sc.age <- as.numeric(as.Date(m.g$statin.sc.date) - as.Date(m.g$dob))/365.25
m.g$statin.bl.age <- as.numeric(as.Date(m.g$statin.bl.date) - as.Date(m.g$dob))/365.25
```

#### Statin bl/sc combined 
```{r}
m.g$statin.sc.bl <- apply(m.g[, c("statin.sc", "statin.bl")], 1, function(x) {
    if(is.na(x[1])) {x[2]}
    else {x[1]}
})
sum(m.g$statin.sc.bl, na.rm = TRUE)
sum(m.g$statin.bl, na.rm = TRUE)
sum(m.g$statin.sc, na.rm = TRUE)
```

## Final data
```{r}
m.final <- select(m.g, 
                  fid = FID, iid = IID, tc.pgs, sex.f = SEX_F, education = EDUCATION, apoe.geno, apoe4.status, fh.ad,
                  sc.smoking, contains("statin"), contains("tc"),
                  bl.age, bl.dx, bl.ad, bl.ci, contains("bl.csf"), bl.atn.a, 
                  lv.age, lv.dx, lv.ad, lv.ci, contains("lv.csf"), lv.atn.a, 
                  contains("conv"),
                  PC1:PC10)
summary(m.final)
```

## Write data
```{r}
write.table(m.final, file = "adni_summary_data.txt", 
            row.names = FALSE, quote = FALSE, sep = "\t")
```