---
title: "CSF abeta42 at baseline ~ TC pgs in ADNI"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 3
    fig_width: 6.5
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input
Relative to project base directory:
```{r}
pheno.file <- "data/adni/adni_summary_data.txt"
```

## Libraries
```{r}
library(tidyverse)
library(gridExtra)

source(paste(path.to.project, "scripts/project_functions.R", sep = ""))
```

## Data
```{r}
pheno <- read.within.project.h(pheno.file)
summary(pheno)
```

### Filter data for presence of CSF data & family history
```{r}
csf <- filter(pheno, !is.na(bl.csf.ab42) & fh.ad == 1 & !is.na(tc.pgs)) %>%
    select(iid, tc.pgs, sex = sex.f, apoe4.status, age = bl.age, statin = statin.sc.bl, 
           bl.dx, csf.ab42 = bl.csf.ab42, tc.mM = sc.tc.mM, PC1:PC10)
summary(csf)
```

Missing data for 31 individuals on statin use and for 3 individuals on TC levels. Filtering individuals with lacking statin data because will be included as interaction factor in analysis.

### Factorize data
```{r}
csf.f <- filter(csf, !is.na(statin))
csf.f$apoe4.status <- factor(csf.f$apoe4.status, levels = c(0, 1), labels = c("APOE4_neg", "APOE4_pos"))
csf.f$sex <- factor(csf.f$sex, levels = c(0, 1), labels = c("Male", "Female"))
csf.f$statin <- factor(csf.f$statin, levels = c(0, 1), labels = c("Statin_neg", "Statin_pos"))

# data for descriptives
m.dt <- filter(csf, !is.na(statin))
m.dt$sex <- factor(m.dt$sex, levels = c(0, 1), labels = c("Male", "Female"))
```

## Descriptives
```{r}
desc.stat <- sum.stat(m.dt, strat_by = "sex", 
         cont_variables = c("age", "tc.mM", "tc.pgs", "csf.ab42"), 
         disc_variables = c("statin", "apoe4.status"))
desc.stat[, " "] <- sapply(desc.stat$p, function(x) if(x <= 0.05) {"*"} else {""})
desc.stat
```

## EDA

**Score & CSF distributions**
```{r, fig.height=6, echo=FALSE}
grid.arrange(
    ggplot(csf.f, aes(tc.pgs)) + 
        geom_histogram(data = filter(csf.f, sex == "Female"), bins = 10, fill = "#1f3c6dff", alpha = 0.5) +
        geom_histogram(data = filter(csf.f, sex == "Male"), bins = 10, fill = "#f5dd4dff", alpha = 0.5) +
        my.theme, 
    
    ggplot(csf.f, aes(tc.pgs, fill = sex)) +
        geom_density(alpha = 0.5) +
        scale_fill_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme,
    
    ggplot(csf.f, aes(csf.ab42)) + 
        geom_histogram(data = filter(csf.f, sex == "Female"), bins = 10, fill = "#1f3c6dff", alpha = 0.5) +
        geom_histogram(data = filter(csf.f, sex == "Male"), bins = 10, fill = "#f5dd4dff", alpha = 0.5) +
        my.theme, 
    
    ggplot(csf.f, aes(csf.ab42, fill = sex)) +
        geom_density(alpha = 0.5) +
        scale_fill_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme,
    
    ncol = 2
)
```

CSF data is clearly skewed. Log transforming for analysis:
```{r}
csf.f$log.csf.ab42 <- log(csf.f$csf.ab42)
```

```{r, echo=FALSE}
grid.arrange(
   ggplot(csf.f, aes(log.csf.ab42)) + 
        geom_histogram(data = filter(csf.f, sex == "Female"), bins = 10, fill = "#1f3c6dff", alpha = 0.5) +
        geom_histogram(data = filter(csf.f, sex == "Male"), bins = 10, fill = "#f5dd4dff", alpha = 0.5) +
        my.theme, 
    
    ggplot(csf.f, aes(log.csf.ab42, fill = sex)) +
        geom_density(alpha = 0.5) +
        scale_fill_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme,
    
    ncol = 2
)
```

### TC pgs effect on CSF data
**Statistics with linear regression**
```{r}
formula.1 <- as.formula("log.csf.ab42 ~ PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 + age + apoe4.status + statin * sex * tc.pgs")
linreg.1 <- lm(formula.1, data = csf.f)
```

**Assumptions:**
```{r, fig.height=6}
par(mfrow = c(2, 2))
plot(linreg.1)
```

**Results:**
```{r}
summary(linreg.1)
```

## Plots
```{r, echo=FALSE}
ggplot(csf.f, aes_string(x = "tc.pgs", y = "log.csf.ab42", color = "sex")) +
    geom_point(size = 3) +
    geom_smooth(method = "lm", se = FALSE) +
    scale_color_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
    facet_grid(~statin) +
    my.theme
```

## Save data
```{r}
save(list = c("csf.f", "desc.stat"), file = "adni_csf_ab.RData")
```

