---
title: "ROSMAP data summarized"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Notes
Covariate data should contain iid, sex, apoe4 status, genetic PC1-PC10, and age. Raw data will be extracted for ROSMAP individuals with European anestry. Statin use and TC levels not available.

## Input
From outside project:
```{r}
m.file <- "~/Research/Cohort_data/ROSMAP/phenotype_data/ROSMAP Other data/ROSMAP_Clinical_Data FINAL.xlsx"
key.file <- "~/Research/Cohort_data/ROSMAP/phenotype_data/ROSMAP Other data/ROSMAP_IDkey.xlsx"
pca.file <- "~/Research/Cohort_data/ROSMAP/genetic_data/pca/PCA-1kG_v3_LD/PCA_EUR/PCA/ROSMAP_EUR_PCA.eigenvec"

# relative to project directory
score.file <- "analyses/rosmap/1_pg_scoring/score_files/roma_tc_pgs.S09.profile"
```

## Libraries
```{r}
library(tidyverse)
library(gridExtra)
library(openxlsx)

source(paste(path.to.project, "scripts/project_functions.R", sep = ""))
source("~/Research/Methods/R Programming/Scripts/Stats/Quantiles.R")
```

## Data

#### Key
```{r}
read.xlsx(key.file) %>%
    colnames()

key <- read.xlsx(key.file)
head(key)
```

#### Genetic PC data
```{r}
pca <- read.table(pca.file, header = TRUE) %>%
    select(gwas_id = IID, PC1:PC10) %>%
    merge.data.frame(select(key, projid, gwas_id)) %>%
    unique()
pca.n <- length(unique(pca$gwas_id))
summary(pca)
```

There are **`r pca.n`** individuals with genetic data & European ancestry. 

#### PGS data
```{r}
pgs <- read.table(paste(path.to.project, score.file, sep = ""), 
                  header = TRUE) %>%
    rename(gwas_id = IID) %>%
    merge.data.frame(select(key, projid, gwas_id)) %>%
    unique()
summary(pgs)
```

**Standardized PGS**
```{r}
pgs$tc.pgs <- scale(pgs$SCORE)
```

**PGS quantiles**
```{r}
pgs$qs5 <- my.quantile.function(pgs$tc.pgs, 5)
```

There are **`r nrow(pgs)`** individuals with European ancestry and PGS data. 

#### Pathology & demographics
```{r}
read.xlsx(m.file) %>% colnames()
pheno <- read.xlsx(m.file) %>%
    merge.data.frame(select(key, projid, gwas_id), all.x = TRUE)
summary(pheno)
```

**APOE-E4 status**
```{r}
my.apoe.f <- function(x) {
    if(!is.na(x)) {as.numeric(grepl("4", x))}
    else {NA}
}

pheno$apoe4.status <- sapply(pheno$apoe_genotype, my.apoe.f)
```

#### Merged data
```{r}
sum(pgs$gwas_id %in% pheno$gwas_id)

m <- merge.data.frame(pgs, pca, all.x = TRUE) %>%
    merge.data.frame(pheno, by = "gwas_id") %>%
    mutate(sex.f = abs(msex - 1), 
           age = as.numeric(gsub("90\\+", 90, age_death)), 
           age.ad = as.numeric(gsub("90\\+", 90, age_first_ad_dx))) %>%
    select(iid = gwas_id, PC1:PC10, sex.f, 
           apoe.geno = apoe_genotype, apoe4.status,  
           age, age.ad, 
           educ, pmi, 
           braaksc, ceradsc, cogdx, 
           tc.pgs, qs5) %>%
    unique()
summary(m)
```

## New variables

#### Simplified cognitive variable

In the 'cogdx' column, 1 is CTRL (no cognitive impairment), 2 & 3 is MCI, and 4 & 5 signifies probable and possible AD respectively. 6 are other type of dementia. This will be simplified into 
- 1 =       CTRL
- 2 & 3 =   MCI
- 4 & 5 =   AD
- 6 =       Other dementia
```{r}
temp <- m$cogdx
temp[temp == 3] <- 2
temp[temp == 4 | temp == 5] <- 3
m$cogdx.2 <- factor(temp, levels = c(1,2,3,6), labels = c("NCI", "MCI", "AD", "Other dementia"))
rm(temp)
```

#### Cognitive binary variables
```{r}
m$cog.ad <- sapply(m$cogdx.2, function(x) {
    if(is.na(x)) {NA}
    else if(x == "NCI") {0}
    else if(x == "AD") {1}
    else {NA}
})

m$cog.ci <- sapply(m$cogdx.2, function(x) {
    if(is.na(x)) {NA}
    else if(x == "NCI") {0}
    else {1}
})
```

#### Pathology in relation to clinical diagnosis
```{r, echo=FALSE}
m.f <- filter(m, !is.na(cogdx.2))
grid.arrange(
    ggplot(m.f, aes(1, fill = factor(ceradsc))) +
        geom_bar(position = "fill") +
        scale_fill_viridis_d(option = "E") +
        coord_polar("y") +
        facet_grid(~cogdx.2) +
        labs(title = "Amyloid pathology") +
        theme(axis.text = element_blank(), 
          axis.title = element_blank(),
          axis.ticks = element_blank()),

    ggplot(m.f, aes(1, fill = factor(braaksc))) +
        geom_bar(position = "fill") +
        scale_fill_viridis_d(direction = -1) +
        coord_polar("y") +
        facet_grid(~cogdx.2) +
        labs(title = "Tau pathology") +
        theme(axis.text = element_blank(), 
          axis.title = element_blank(),
          axis.ticks = element_blank()),
    
    ggplot(filter(m, ceradsc == 1 | ceradsc == 2), aes(1, fill = factor(braaksc))) +
        geom_bar(position = "fill") +
        scale_fill_viridis_d(direction = -1) +
        coord_polar("y") +
        facet_grid(~cogdx.2) +
        labs(title = "Tau pathology in CERAD positive individuals") +
        theme(axis.text = element_blank(), 
              axis.title = element_blank(),
              axis.ticks = element_blank()),
    nrow = 3
)
```

#### Binary CERAD scores
**CERAD binary**: 1&2 = POS, 3&4 = NEG
```{r}
temp <- m$ceradsc 
temp[temp == 4] <- 3
temp[temp == 1] <- 2
m$cerad.bin <- abs(temp - 3)
```

**CERAD neg vs possible AD**: 3 = POS, 4 = NEG
```{r}
temp <- abs(m$ceradsc - 4)

cerad.poss.ad <- temp
cerad.poss.ad[cerad.poss.ad %in% c(2:3)] <- NA

m$cerad.poss.ad <- cerad.poss.ad
```

**CERAD neg vs probable AD**: 2 = POS, 4 = NEG
```{r}
cerad.prob.ad <- temp
cerad.prob.ad[cerad.prob.ad %in% c(1,3)] <- NA
cerad.prob.ad[cerad.prob.ad == 2] <- 1

m$cerad.prob.ad <- cerad.prob.ad
```

**CERAD neg vs definite AD**: 1 = POS, 4 = NEG
```{r}
cerad.def.ad <- temp
cerad.def.ad[cerad.def.ad %in% c(1:2)] <- NA
cerad.def.ad[cerad.def.ad == 3] <- 1

m$cerad.def.ad <- cerad.def.ad
```

**CERAD neg vs all AD**: 1 = POS, 4 = NEG
```{r}
all.ad <- temp
all.ad[all.ad %in% c(2,3)] <- 1

m$cerad.all.ad <- all.ad
```

#### Biological AD
Biological AD defined as a CERAD score of 1 or 2 and a clinical diagnosis of 4 or 5. Controls are defined as CERAD score 3 or 4 and a cognitive score of 1. 
```{r}
m$adbiol <- apply(m[, c("cerad.bin", "cogdx")], 1, function(y) {
    x <- as.character(y)
    if(is.na(x[1]) | is.na(x[2])) {NA}
    else if(x[1] == "1" & (x[2] == "4" | x[2] == "5")) {1}
    else if(x[1] == "0" & x[2] == "1") {0}
    else {NA}
    }
    ) %>%
    as.character() %>%
    as.numeric()
```

Number of subjects for biological AD:
```{r}
group_by(m, adbiol) %>%
    summarize(n = length(adbiol))
```

## Final data
```{r}
summary(m)
```

## Write data
```{r}
write.table(m, file = "rosmap_summary_data.txt", 
            quote = FALSE, row.names = FALSE, sep = "\t")
```