---
title: "PET tau ~ TC-PGS in PRAD"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 3
    fig_width: 6.5
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Notes
PREVENT-AD have a low of individuals and a low number of individuals on statins. Thus, in order to preserve power, and the fact that the sex effect of the TC-PGS on TC levels were pronounced for both PREVENT-AD and ADNI (whereas stratification on statins in ADNI had smaller effect), we are examining the effect of the interaction between TC-PGS and sex on PET tau. Statin use is included as a covariate. 

## Input
Relative to project base directory:
```{r}
pheno.file <- "data/preventad/prad_summary_data.txt"
```

## Libraries
```{r}
library(tidyverse)
library(gridExtra)

source(paste(path.to.project, "scripts/project_functions.R", sep = ""))
```

## Data
```{r}
pheno <- read.within.project.h(pheno.file)
summary(pheno)
```

### Cleaning data
```{r}
pet <- filter(pheno, !is.na(pettau.etx)) %>%
    select(iid, PC1:PC10, sex, apoe4.status, statin, tc.mM, pettau.age, pettau.etx, pettau.pos, tc.pgs) %>%
    mutate(sex = relevel(sex, ref = "Male"))
```

**Missing data**
```{r}
apply(pet, 2, function(x) sum(is.na(x)))
filter(pet, is.na(statin) | is.na(tc.mM) | is.na(pettau.age | is.na(sex))) %>%
    select(iid, sex:tc.pgs)
```
Three individuals have missing data (either on statin, TC levels or age). TC levels are only for descriptives why only "MTL0143" will be excluded. 

```{r}
pet.cc <- filter(pet, iid != "MTL0143")
```

## Descriptives
```{r}
# converting binary factors to numeric (0 or 1)
m.dt <- pet.cc
m.dt$apoe4.status <- as.numeric(m.dt$apoe4.status) - 1
m.dt$statin <- as.numeric(m.dt$statin) - 1
m.dt$pettau.pos <- as.numeric(m.dt$pettau.pos) - 1

desc.stat <- sum.stat(m.dt, strat_by = "sex", 
         cont_variables = c("pettau.age", "tc.mM", "tc.pgs", "pettau.etx"), 
         disc_variables = c("statin", "apoe4.status", "pettau.pos"))
desc.stat[, " "] <- sapply(desc.stat$p, function(x) if(x <= 0.05) {"*"} else {""})
desc.stat
```

## EDA

#### TC-PGS distribution
```{r}
grid.arrange(
    ggplot(pet.cc, aes(tc.pgs)) + 
        geom_histogram(data = filter(pet.cc, sex == "Female"), bins = 10, fill = "#f5dd4dff") +
        geom_histogram(data = filter(pet.cc, sex == "Male"), bins = 10, fill = "#1f3c6dff") +
        my.theme, 
    
    ggplot(pet.cc, aes(tc.pgs, fill = sex)) +
        geom_density(alpha = 0.5) +
        scale_fill_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme,
    
    ncol = 2
)
```

#### PET distribution
```{r}
grid.arrange(
    ggplot(pet.cc, aes(pettau.etx)) + 
        geom_histogram(data = filter(pet.cc, sex == "Female"), bins = 10, fill = "#f5dd4dff") +
        geom_histogram(data = filter(pet.cc, sex == "Male"), bins = 10, fill = "#1f3c6dff") +
        my.theme, 
    
    ggplot(pet.cc, aes(log2(pettau.etx), fill = sex)) +
        geom_density(alpha = 0.5) +
        scale_fill_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme,
    
    ncol = 2
)
```

Data look fairly normally distributed. 

## TC-PGS effect on PET data

### Linear regression - All individuals

**Statistics**
```{r}
formula.1 <- as.formula("log2(pettau.etx) ~ PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10+pettau.age + apoe4.status + statin + sex * tc.pgs")
linreg.1 <- lm(formula.1, data = pet.cc)
```

Assumptions:
```{r, fig.height=6}
par(mfrow = c(2, 2))
plot(linreg.1)
```

Results:
```{r}
summary(linreg.1)
```

```{r}
formula.2 <- as.formula("pettau.pos ~ PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10+pettau.age + apoe4.status + statin + sex * tc.pgs")
logreg <- glm(formula.2, data = pet.cc, family = "binomial")
summary(logreg)
```

**Plotting**
```{r, echo=FALSE, fig.width=3.25}
ggplot(pet.cc, aes_string(x = "tc.pgs", y = "pettau.etx", color = "sex")) +
        geom_point(size = 3) +
        geom_smooth(method = "lm", se = FALSE) +
    scale_y_continuous(trans = "log2", 
                       breaks = c(1, 1.25, 1.5)) +
        scale_color_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme +
    theme(axis.ticks.y =)
```

## Save data
```{r}
save(list = c("pet.cc", "desc.stat"), file = "prad_pet_tau.RData")
```

