---
title: "Hypercholesterolemia ROC in PRAD"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Notes
The effect of a tc-pgs (with SNPs with p <= 1e-6) on a hypercholesterolemia diagnosis in PREVENT-AD. Due to the small sample size, individuals were judged to be hypercholesterolemic if they were on statins or if they had TC levels > 6.2 mM (according to Mayo clinic). 

## Input
Relative to project base directory:
```{r}
pheno.file <- "data/preventad/prad_summary_data.txt"
```

## Libraries
```{r}
library(tidyverse)
library(gridExtra)
library(pROC)

source(paste(path.to.project, "scripts/project_functions.R", sep = ""))
```

## Data
```{r}
pheno <- read.within.project.h(pheno.file) %>%
    mutate(tc.age2 = tc.age^2) %>%
    select(iid, PC1:PC10, sex, statin, tc.age, tc.age2, apoe4.status, tc.pgs, tc.mM) %>%
    filter(!is.na(tc.mM) & !is.na(apoe4.status) & !is.na(tc.pgs))
summary(pheno)
```

### Hypercholesterolemia variables
Two variables are created, "hc.tc" which is defined as tc.mM > 6.2, and "hc.statin" which is defined as tc.mM > 6.2 in statin negative individuals, and includes all statin users as hypercholesterolemic. 
```{r}
pheno$hc.tc <- as.numeric(pheno$tc.mM > 6.2)

pheno$hc.statin <- apply(pheno[, c("statin", "tc.mM")], 1, function(x) {
    if(is.na(x[1])) {NA}
    else if(x[1] == "Statin_pos") {1}
    else if(x[1] == "Statin_neg" & x[2] > 6.2) {1}
    else if(x[1] == "Statin_neg" & x[2] <= 6.2) {0}
})
```

### Summary statistics

Discrete variables are presented as percentages and continuous variables as means. Difference in continuous and discrete variables between strata is assessed by Welch t-test and chi-squared test, respectively.
```{r}
# converting binary factors to numeric (0 or 1)
m.dt <- pheno
m.dt$apoe4.status <- as.numeric(m.dt$apoe4.status) - 1

desc.stat <- sum.stat(m.dt, strat_by = "sex", 
         cont_variables = c("tc.age", "tc.mM"), 
         disc_variables = c("apoe4.status", "hc.statin"))
desc.stat[, " "] <- sapply(desc.stat$p, function(x) if(x <= 0.05) {"*"} else {""})
desc.stat
```

## Replicate R2 values from bootstrap analysis
```{r}
statin.neg <- filter(pheno, statin == "Statin_neg")
rep.cov <- lm(tc.mM ~ PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10+apoe4.status+tc.age+tc.age2+sex, 
              data = statin.neg)
statin.neg$tc.res <- rep.cov$residuals

rep.full <- lm(tc.mM ~ PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10+apoe4.status+tc.age+tc.age2+sex+tc.pgs, 
              data = statin.neg)
rep.res <- (lm(tc.res ~ tc.pgs, data = statin.neg)) 

summary(rep.cov)$adj.r.squared
summary(rep.full)$adj.r.squared
summary(rep.res)$adj.r.squared
summary(rep.full)$adj.r.squared-summary(rep.cov)$adj.r.squared
```

```{r}
statin.neg.f <- filter(statin.neg, sex == "Female")

sex.cov <- lm(tc.mM ~ PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10+apoe4.status+tc.age+tc.age2, 
              data = statin.neg.f)
statin.neg.f$tc.res <- sex.cov$residuals

sex.full <- lm(tc.mM ~ PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10+apoe4.status+tc.age+tc.age2+tc.pgs, 
              data = statin.neg.f)

sex.res <- (lm(tc.res ~ tc.pgs, data = statin.neg.f)) 

summary(sex.cov)$adj.r.squared
summary(sex.full)$adj.r.squared
summary(sex.res)$adj.r.squared
summary(sex.full)$adj.r.squared-summary(sex.cov)$adj.r.squared
```

## ROC

### Females
```{r}
# Females
m.f <- filter(pheno, sex == "Female" & !is.na(hc.statin))

my.log1 <- glm(hc.statin~PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10+tc.age+tc.age2, data = m.f, family = "binomial")
my.log2 <- glm(hc.statin~PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10+tc.age+tc.age2+tc.pgs, data = m.f, family = "binomial")

m.f$prob1 <- predict(my.log1, type = "response")
m.f$prob2 <- predict(my.log2, type = "response")

g <- roc(hc.statin ~ prob1, data = m.f)
h <- roc(hc.statin ~ prob2, data = m.f)

roc.test(g, h)
```

### Males
```{r}
m.m <- filter(pheno, sex == "Male" & !is.na(hc.statin))

my.log3 <- glm(hc.statin~PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10+tc.age+tc.age2, data = m.m, family = "binomial")
my.log4 <- glm(hc.statin~PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10+tc.age+tc.age2+tc.pgs, data = m.m, family = "binomial")

m.m$prob3 <- predict(my.log3, type = "response")
m.m$prob4 <- predict(my.log4, type = "response")

j <- roc(hc.statin ~ prob3, data = m.m)
k <- roc(hc.statin ~ prob4, data = m.m)

roc.test(j, k)
```

## Plots
```{r, echo = FALSE, fig.height=4.5}
ggroc(list("F; covariates" = g, "F; covariates + TC-PGS" = h, 
           "M; covariates" = j, "M; covariates + TC_PGS" = k), size = 1) +
        scale_color_viridis_d(option = "E") +
        annotate(x = 0, y = 0, 
                 geom = "text", 
                 label = paste("AUC:\nF; covariates: ", round(g$auc, 3), 
                           "\nF; covariates + PGS: ",round(h$auc, 3),
                           "\nM; covariates: ", sprintf("%.3f", round(j$auc, 3)), 
                           "\nM; covariates + PGS: ", round(k$auc, 3), sep = ""), 
                 hjust = 1, vjust = 0) +
         my.theme +
    theme(legend.position = "right")
```

## Save data
```{r}
save(list = c("g", "h", "j", "k"), file = "prad_hypchol.RData")
```

