---
title: "adni_tc_boot_data"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input
Relative to project base directory:
```{r}
mydata.file <- "data/adni/adni_summary_data.txt"
```

Output parameters:
```{r}
output.name <- "adni_tc_boot_data"
```

## Libraries
```{r}
library(tidyverse)

source(paste(path.to.project, "scripts/project_functions.R", sep = ""))
```

## Read & select data
```{r}
mydata <- read.within.project.h(mydata.file)
summary(mydata)
```

Number of data points after stratification for family history, statin use and sex:
```{r}
filter(mydata, fh.ad == 1) %>%
    group_by(fh.ad, statin.sc.bl, sex.f) %>%
    summarise(n.tc = sum(!is.na(sc.tc.mM)), 
              n.ctrl = sum(bl.dx == 1), 
              n.mci = sum(bl.dx == 2), 
              n.ad = sum(bl.dx == 3))
```


```{r}
mydata.s <- filter(mydata, fh.ad == 1 & !is.na(statin.sc.bl) & !is.na(sc.tc.mM)) %>% 
    select(iid, PC1:PC10, sex = sex.f, apoe4.status, statin = statin.sc.bl, sc.smoking, age = tc.age, bl.dx, tc.mM = sc.tc.mM) %>%
    mutate(age2 = age^2) # create squared age to include as covariate
summary(mydata.s)
```

Missing data:
```{r}
apply(mydata.s, 2, function(x) sum(is.na(x)))
```

## Covariates relationship with TC levels
Each covariate evaluated separately by linear regression with TC levels. Resulting p-values:
```{r}
sapply(c("sex", "apoe4.status", "statin", "sc.smoking", "age", "age2", "bl.dx"), function(x) {
    summary(lm(as.formula(paste("tc.mM ~", x)), data = mydata.s))$coefficients[2,4]
})
```

## Write data
```{r}
my.write.t(mydata.s, file.name = paste(output.name, ".txt", sep = ""))
```
