---
title: "Bootstrap TC ~ TC pgs in PRAD"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 4.5
    fig_width: 11
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

Score files, relative to project directory: 
```{r}
score.files <- "analyses/adni/v1_ownclumping/1_pg_scoring/score_files/" # folder with score files
score.range <- "analyses/adni/v1_ownclumping/1_pg_scoring/scoring_range.txt" # scoring range file
```

Phenotype data:
```{r}
pheno.file <- "adni_tc_boot_data.txt"

dep.var <- "tc.mM" # colname of dependent variable
covars <- c("sex", "age", "age2", "apoe4.status") # character vector with colnames of covariates (PCs will be added separately)
int.var <- "statin" # variable to split on
```

Analysis parameters:
```{r}
n.pcs <- 10 # number of genetic principal components to include as covariates
n.bootstrap <- 5000
```

Output parameters:
```{r}
output.name.org <- "adni_tc_boot_statinsplit"
```

## Libraries
```{r}
library(tidyverse)
library(rcompanion)
library(boot)
library(data.table)

source(paste(path.to.project, "scripts/project_functions.R", sep = ""))
```

## Data
```{r}
pheno <- read.table(pheno.file, header = TRUE, sep = "\t") %>%
    rename(IID = iid)
head(pheno)

pcs <- paste("PC", 1:n.pcs, sep = "")
```

**Summary statistics**
Discrete variables are presented as percentages and continuous variables as means. Difference in continuous and discrete variables between strata is assessed by Welch t-test and chi-squared test, respectively.
```{r}
# converting binary factors to numeric (0 or 1)
m.dt <- pheno
m.dt$statin <- factor(m.dt$statin, levels = c(0, 1), labels = c("Statin_neg", "Statin_pos"))

desc.stat <- sum.stat(m.dt, strat_by = int.var, 
         cont_variables = c("age", "tc.mM"), 
         disc_variables = c("sex", "apoe4.status"))
desc.stat[, " "] <- sapply(desc.stat$p, function(x) if(x <= 0.05) {"*"} else {""})
desc.stat
```

## PG scoring
Strata levels:
```{r}
pheno$statin <- factor(pheno$statin, levels = c(0,1), labels = c("Statin_neg", "Statin_pos"))
str.1 <- levels(pheno[, int.var])[1]
str.2 <- levels(pheno[, int.var])[2]
```

```{r}
# New complete paths ####
score.files <- paste(path.to.project, score.files, sep = "/")
score.range <- paste(path.to.project, score.range, sep = "/")

output.dir      <- paste(path.to.project, path.to.wd, sep = "")

my.ls1 <- c(ls(), "my.ls1")
my.ls1
```

**All subjects**
```{r}
m.cc <- pheno
output.name <- output.name.org
covars <- c(covars, int.var)

source(paste(path.to.project, "scripts/lm_boot_forRmd.R", sep = ""))
```

**`r str.1` subjects**
```{r}
ls2 <- ls()
rm(list = ls2[!ls2 %in% my.ls1])

m.cc <- filter(pheno, (!!as.symbol(int.var)) == str.1)
output.name <- paste(output.name.org, str.1, sep = "_")
covars <- covars[!grepl(int.var, covars)]

source(paste(path.to.project, "scripts/lm_boot_forRmd.R", sep = ""))
```

**`r str.2` subjects**
```{r}
ls2 <- ls()
rm(list = ls2[!ls2 %in% my.ls1])

m.cc <- filter(pheno, (!!as.symbol(int.var)) == str.2)
output.name <- paste(output.name.org, str.2, sep = "_")
covars <- covars[!grepl(int.var, covars)]

source(paste(path.to.project, "scripts/lm_boot_forRmd.R", sep = ""))
```

**Number of SNPs per score**
```{r}
select(m.cc, contains("CNTS")) %>%
    unique() %>%
    apply(2, unique)
```

## Results plotted
```{r fig1, echo=FALSE}
all <- read.table(paste(output.name.org, "Results.txt", sep = "_"), header = TRUE) %>%
  mutate(pop = "all")
neg <- read.table(paste(output.name.org, str.1, "Results.txt", sep = "_"), header = TRUE)  %>%
  mutate(pop = str.1)
pos <- read.table(paste(output.name.org, str.2, "Results.txt", sep = "_"), header = TRUE) %>%
  mutate(pop = str.2)

m2 <- rbind.data.frame(all, neg, pos) %>%
  merge.data.frame(my.range.list)

ggplot(m2, aes(x = factor(pmax), y = Original.value.R2, color = -log10(Original.value.P))) +
    geom_point(size = 3, shape = 18) +
    geom_errorbar(aes(ymin = Original.value.R2 - SD.value.R2, ymax = Original.value.R2 + SD.value.R2), 
                  width = 0.3, size = 1) +
    labs(title = paste(dep.var, "~ TC-PGS"), 
         caption = paste("corrected for:", paste(c(paste("PC1-", n.pcs, sep = ""), covars, "and", int.var, "for the 'all' group"), collapse = ", "), "\nerror bars indicate SD"), 
         y = "Additional R2 explained by PGS", 
         x = "P cut-off") +
    scale_color_viridis_c(name = "-log10(p)", 
                          option = "E", direction = 1) +
    my.theme +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
          legend.position = "right") +
    facet_grid(~pop)
```

## Save data
```{r}
save(list = "m2", file = paste(output.name.org, ".RData", sep = ""))
```

