---
title: "PET amyloid at baseline ~ TC pgs in PRAD"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 3
    fig_width: 6.5
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Notes
PREVENT-AD have a low of individuals and a low number of individuals on statins. Thus, in order to preserve power, and the fact that the sex effect of the TC-PGS on TC levels were pronounced for both PREVENT-AD and ADNI (whereas stratification on statins in ADNI had smaller effect), we are examining the effect of the interaction between TC-PGS and sex on CSF amyloid. Statin use is included as a covariate. 

Same as v1 but added combined analysis with robust regression.

## Input
Relative to project base directory:
```{r}
pheno.file <- "data/preventad/prad_summary_data.txt"
```

## Libraries
```{r}
library(tidyverse)
library(gridExtra)

source(paste(path.to.project, "scripts/project_functions.R", sep = ""))
```

## Data
```{r}
pheno <- read.within.project.h(pheno.file)
summary(pheno)
```

### Cleaning data
```{r}
pet <- filter(pheno, !is.na(petab)) %>%
    mutate(petab.pos = as.numeric(petab > 1.37)) %>%
    select(iid, PC1:PC10, sex, apoe4.status, statin, tc.mM, petab.age, petab, petab.pos, tc.pgs)
```

**Missing data**
```{r}
apply(pet, 2, function(x) sum(is.na(x)))
filter(pet, is.na(statin) | is.na(tc.mM)) %>%
    select(iid, sex:tc.pgs)
```
Two individuals have missing data (either on statin or TC levels). Due the small sample size, both are kept. 

```{r}
pet.cc <- pet
```

## Descriptives
```{r}
# converting binary factors to numeric (0 or 1)
m.dt <- pet.cc
m.dt$apoe4.status <- as.numeric(m.dt$apoe4.status) - 1
m.dt$statin <- as.numeric(m.dt$statin) - 1

desc.stat <- sum.stat(m.dt, strat_by = "sex", 
         cont_variables = c("petab.age", "tc.mM", "tc.pgs", "petab"), 
         disc_variables = c("statin", "apoe4.status"))
desc.stat[, " "] <- sapply(desc.stat$p, function(x) if(x <= 0.05) {"*"} else {""})
desc.stat
```

## EDA

**Score distribution**
```{r, echo=FALSE, fig.height=6}
grid.arrange(
    ggplot(pet.cc, aes(tc.pgs)) + 
        geom_histogram(data = filter(pet, sex == "Female"), bins = 10, fill = "#1f3c6dff") +
        geom_histogram(data = filter(pet, sex == "Male"), bins = 10, fill = "#f5dd4dff") +
        my.theme, 
    
    ggplot(pet.cc, aes(tc.pgs, fill = sex)) +
        geom_density(alpha = 0.5) +
        scale_fill_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme,

    ggplot(pet.cc, aes(petab)) + 
        geom_histogram(data = filter(pet, sex == "Female"), bins = 10, fill = "#1f3c6dff") +
        geom_histogram(data = filter(pet, sex == "Male"), bins = 10, fill = "#f5dd4dff") +
        my.theme, 
    
    ggplot(pet.cc, aes(petab, fill = sex)) +
        geom_density(alpha = 0.5) +
        scale_fill_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme,
    
    ncol = 2
)
```

PET AB is not normally distributed. However, these values could still be biologically relevant; PiB index SUVR values of 3 is definitely possible (see Mormino et al., 2008, https://doi.org/10.1093/brain/awn320). Thus, data can be grouped on amyloid positivity or analyzed with robust regression. 

**Amyloid positivity**
```{r, fig.width=3.25}
ggplot(pet.cc, aes(petab, fill = factor(petab.pos))) +
        geom_density(alpha = 0.5) +
        scale_fill_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme
```


```{r}
pet.cc %>%
    group_by(petab.pos, sex) %>%
    summarise(n = sum(!is.na(petab)))
```

### TC pgs effect on PET data
Due to the non-normally distributed data, the data are analyzed in three ways, first with linear regressions stratified for amyloid positivity, and secondly with a logistic regression with amyloid positivity as the dependent variable and thirdly by robust regression. 

#### Amyloid negative individuals; linear regression
```{r}
formula.1 <- as.formula("petab ~ PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 + petab.age + apoe4.status + statin + sex * tc.pgs")
linreg.1 <- lm(formula.1, data = filter(pet.cc, petab.pos == 0))
```

**Assumptions:**
```{r, fig.height=6}
par(mfrow = c(2, 2))
plot(linreg.1)
```

**Results:**
```{r}
summary(linreg.1)
```

#### Amyloid positive individuals; linear regression
There is only 18 individuals that are considered amyloid positive, thus adjusting the the model, excluding any interaction terms and the PCs.
```{r}
formula.2 <- as.formula("petab ~ petab.age + apoe4.status + statin + sex + tc.pgs")
linreg.2 <- lm(formula.2, data = filter(pet.cc, petab.pos == 1))
```

**Assumptions:**
```{r, fig.height=6}
par(mfrow = c(2, 2))
plot(linreg.2)
```

**Results:**
```{r}
summary(linreg.2)
```

#### Amyloid positivity; logistic regression
```{r}
formula.3 <- as.formula("petab.pos ~ PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 + petab.age + apoe4.status + statin + sex * tc.pgs")
logreg <- glm(formula.3, data = pet.cc, family = "binomial")
summary(logreg)
```

**Odds ratios**
```{r}
exp(coef(logreg))
```

#### Robust regression
```{r}
library(MASS)
library(sfsmisc)
```

```{r}
rr.f <- as.formula("petab ~ PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 + petab.age + apoe4.status + statin + sex * tc.pgs")
rr.m <- rlm(rr.f, data = pet.cc, psi = "psi.bisquare")
summary(rr.m)
```

```{r}
f.robftest(rr.m, "tc.pgs")
```

```{r}
f.robftest(rr.m, "sexMale:tc.pgs")
```

## Plots
```{r, echo=FALSE}
grid.arrange(
    ggplot(filter(pet.cc, petab.pos == 0), aes_string(x = "tc.pgs", y = "petab", color = "sex")) +
        geom_point(size = 3) +
        geom_smooth(method = "lm", se = FALSE) +
        scale_color_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        labs(title = "Amyloid negative") +
        my.theme, 
    
    ggplot(filter(pet.cc, petab.pos == 1), aes_string(x = "tc.pgs", y = "petab")) +
        geom_point(size = 3) +
        geom_smooth(method = "lm", se = FALSE) +
        scale_color_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        labs(title = "Amyloid positive") +
        my.theme, 
    
    ggplot(pet.cc, aes_string(x="tc.pgs", y="petab", color="sex")) +
        geom_point(size = 3) +
        geom_smooth(method = "lm", se = FALSE) +
        scale_color_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        labs(title = "Robust regression") +
        my.theme, 
    
    ncol = 3
)

```
