---
title: "CSF amyloid beta pathology at baseline ~ TC pgs in PRAD"
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 3
    fig_width: 6.5
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Notes
Findings from rs72633963 and rs3846662 analyses suggest interaction with APOE-e4 status and results from Kobe et al., 2020, suggest that cholesterol only associate with amyloid pathology in individuals not on vascular medication. Thus we wanted to see whether the TC-PGS would interact with APOE-e4 status in statin free individuals. 

## Input
Relative to project base directory:
```{r}
pheno.file <- "data/preventad/prad_summary_data.txt"
```

## Libraries
```{r}
library(tidyverse)
library(gridExtra)

source(paste(path.to.project, "scripts/project_functions.R", sep = ""))
```

## Data
```{r}
pheno <- read.within.project.h(pheno.file)
summary(pheno)
```

### Cleaning data
```{r}
csf <- filter(pheno, !is.na(csf.ab42)) %>%
    select(iid, PC1:PC10, sex, apoe4.status, statin, tc.mM, csf.age, csf.ab42, tc.pgs)

csf$sex <- factor(csf$sex, levels = c("Male", "Female"))
```

**Missing data**
```{r}
apply(csf, 2, function(x) sum(is.na(x)))
```

Data is missing for APOE4 status and for TC levels, each for one individual. Excluding individual with no APOE4 status since this variable will be incliuded as a covariate, and keeping individual with no TC data because this will only be used for descriptives. 
```{r}
csf.cc <- filter(csf, !is.na(apoe4.status))
```


## Descriptives
```{r}
# converting binary factors to numeric (0 or 1)
m.dt <- csf.cc
m.dt$apoe4.status <- as.numeric(m.dt$apoe4.status) - 1

# descriptives
desc.stat <- sum.stat(m.dt, strat_by = "sex", 
         cont_variables = c("csf.age", "tc.mM", "tc.pgs", "csf.ab42"), 
         disc_variables = c("apoe4.status"))
desc.stat[, " "] <- sapply(desc.stat$p, function(x) if(x <= 0.05) {"*"} else {""})
desc.stat
```

Males and females vary only with regards to TC levels, female having higher levels. 

## EDA

**Score distribution**
```{r, fig.height=9}
grid.arrange(
    ggplot(csf.cc, aes(tc.pgs)) + 
        geom_histogram(data = filter(csf.cc, sex == "Female"), bins = 10, fill = "#f5dd4dff") +
        geom_histogram(data = filter(csf.cc, sex == "Male"), bins = 10, fill = "#1f3c6dff") +
        my.theme, 
    
    ggplot(csf.cc, aes(tc.pgs, fill = sex)) +
        geom_density(alpha = 0.5) +
        scale_fill_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme,
    
    ggplot(csf.cc, aes(csf.ab42)) + 
        geom_histogram(data = filter(csf.cc, sex == "Female"), bins = 10, fill = "#f5dd4dff") +
        geom_histogram(data = filter(csf.cc, sex == "Male"), bins = 10, fill = "#1f3c6dff") +
        my.theme, 
    
    ggplot(csf.cc, aes(csf.ab42, fill = sex)) +
        geom_density(alpha = 0.5) +
        scale_fill_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme,
    
    ncol = 2
)
```

## TC pgs effect on CSF data

### Linear regression - All individuals

**Statistics**
```{r}
formula.1 <- as.formula("csf.ab42 ~ PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 + csf.age + sex + apoe4.status * tc.pgs")
linreg.1 <- lm(formula.1, data = csf.cc)
```

Assumptions:
```{r, fig.height=6}
par(mfrow = c(2, 2))
plot(linreg.1)
```

Results:
```{r}
summary(linreg.1)
```

**Plotting**
```{r, echo=FALSE, fig.width=3.25}
ggplot(csf.cc, aes_string(x = "tc.pgs", y = "csf.ab42", color = "apoe4.status")) +
        geom_point(size = 3) +
        geom_smooth(method = "lm", se = FALSE) +
        scale_color_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme
```

### Linear regression - Statin free individuals

```{r}
csf.statinfree <- filter(csf, statin == "Statin_neg")
```

**Statistics**
```{r}
formula.2 <- as.formula("csf.ab42 ~ PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 + csf.age + sex + apoe4.status * tc.pgs")
linreg.2 <- lm(formula.2, data = csf.statinfree)
```

Assumptions:
```{r, fig.height=6}
par(mfrow = c(2, 2))
plot(linreg.2)
```

Results:
```{r}
summary(linreg.2)
```

**Plotting**
```{r, echo=FALSE, fig.width=3.25}
ggplot(csf.statinfree, aes_string(x = "tc.pgs", y = "csf.ab42", color = "apoe4.status")) +
        geom_point(size = 3) +
        geom_smooth(method = "lm", se = FALSE) +
        scale_color_manual(values = c("#1f3c6dff", "#f5dd4dff")) +
        my.theme
```
