---
title: "Log of 'prad_summary_data.txt'" 
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

#### Cleaned raw data files:
```{r}
demog.file <- "extracted_from_rawdata/prad_demog_data.txt"
statin.file <- "extracted_from_rawdata/prad_statin_data.txt"
tc.file <- "extracted_from_rawdata/prad_lipid_data.txt"
csf.file <- "extracted_from_rawdata/prad_csf_data.txt"
pet.file <- "extracted_from_rawdata/prad_pet_data.txt"
bl.file <- "extracted_from_rawdata/prad_bl_data.txt"
```

#### PCA file from outside project:
```{r}
pca.file <- "~/Research/Cohort_data/ADPREVENT/Genetic_data/PRAD_2019/pca-1kG_v3_ld/PCA_EUR/PCA/PRAD_2019_EUR_PCA.eigenvec"
```

#### TC-PGS file from within project:
p-value cut-off determined to be p <= 1e-6
```{r}
pgs.file <- "analyses/preventad/1_pg_scoring/score_files/prad_tc_pgs.S09.profile"
```

## Libraries
```{r}
library(tidyverse)
library(lubridate)
```

## Data 

#### Genetic PC data
```{r}
pca <- read.table(pca.file, header = TRUE) %>%
    select(iid = IID, PC1:PC10)
head(pca)
```

#### Baseline age
```{r}
bl <- read.table(bl.file, header = TRUE) %>%
    rename(bl.visit = visit)
head(bl)
```

#### Demographic data
```{r}
demog <- read.table(demog.file, header = TRUE)
summary(demog)
```

#### Statin data
```{r}
statin <- read.table(statin.file, header = TRUE) 
summary(statin)
```

#### Lipid data
```{r}
tc <- read.table(tc.file, header = TRUE) 
summary(tc)

tc.preel <- filter(tc, visit == "PREEL00") %>%
    unique() %>% 
    select(iid, tc.age, tc.date, tc.mM) %>%
    filter(!is.na(tc.mM))
```

#### CSF data
```{r}
csf <- read.table(csf.file, header = TRUE)
summary(csf)

csf.bl <- filter(csf, grepl("00", visit)) %>%
    rename(csf.visit = visit)
```

#### PET data
```{r}
pet <- read.table(pet.file, header = TRUE, sep = "\t") %>%
    select(-visit)
summary(pet)
```

#### TC-PGS data
```{r}
pgs <- read.table(paste(path.to.project, pgs.file, sep = ""), header = TRUE)
summary(pgs)
```
No. of SNPs included in the score = `r unique(pgs$CNT)/2`
PGS range = `r range(pgs$SCORE)`

Standardizing the PGS:
```{r}
pgs$tc.pgs <- scale(pgs$SCORE)[,1]
pgs$uw.tcpgs <- scale((pgs$CNT2/pgs$CNT))[,1]

pgs.s <- select(pgs, iid = IID, tc.pgs, uw.tcpgs)
```

#### Merge data
```{r}
m.bl <- Reduce(function(x, y) merge.data.frame(x, y, all.x=TRUE), list(pca, demog, bl, statin, tc.preel, csf.bl, pet, pgs.s))
```

## New variables

#### Date of birth
Many individuals lack data for date of birth but have dates for and age for other data collection. Thus missing date of birth can be calculated with these variables:
```{r}
m.date <- select(m.bl, dob, contains("date"), contains("age")) %>%
    filter(is.na(dob)) %>%
    arrange(tc.age, bl.age, csf.age)
m.date

dob.f <- function(x.date, x.age) {
    date(date(x.date) - dyears(x.age))
}


m.bl <- mutate(m.bl,
                 bl.age.y = bl.age/12,
                 bl.dob.calc = dob.f(bl.date, bl.age.y), 
                 tc.dob.calc = dob.f(tc.date, tc.age), 
                 csf.dob.calc = dob.f(csf.date, csf.age))
m.bl$dob.final <- apply(m.bl[, c("dob", "bl.dob.calc", "tc.dob.calc", "csf.dob.calc")], 1, function(x) {
    x[min(which(!is.na(x)))] 
})

ggplot(m.bl, aes(date(dob), bl.dob.calc)) +
    geom_point()
```

Number of individuals with calculated DOBs:
```{r}
sum(is.na(m.bl$dob)) - sum(is.na(m.bl$dob.final))
```

PET data are collected over different visits for different subjects, thus, age for these visits needs to be calculated.
```{r}
m.bl$petab.age <- as.duration(m.bl$dob.final %--% m.bl$petab.date) / dyears(1)
m.bl$pettau.age <- as.duration(m.bl$dob.final %--% m.bl$pettau.date) / dyears(1)

summary(m.bl)
```

## Check variables classes
```{r}
lapply(m.bl, class)
```

```{r}
m.bl$apoe4.status <- factor(m.bl$apoe4.status, levels = c(0, 1), labels = c("APOE4_neg", "APOE4_pos"))
m.bl$statin <- factor(m.bl$statin, levels = c(0, 1), labels = c("Statin_neg", "Statin_pos"))
m.bl$pettau.pos <- factor(m.bl$pettau.pos, levels = c(0, 1), labels = c("TAU_neg", "TAU_pos"))
```

## Write data
```{r}
write.table(m.bl, file = "prad_summary_data.txt", 
            quote = FALSE, row.names = FALSE, sep = "\t")
```

