---
title: "Log of 'prad_demog_data.txt'" 
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

### Raw data source
```{r, echo=FALSE}
raw.dir <- "~/Research/Cohort_data/ADPREVENT/Patient_data/DATA_RELEASE_5.0_20171130/raw_data/dataDump01Dec17/nn_xlsx/" 
```
`r raw.dir`

Note that data are  collected from PREVENT-AD data release 5.0.

### Raw data files
Paths relative to raw data directory (files and names are exactly the same as raw data files, just converted to '.xlsx' format):
```{r}
info.file <- "candidate_info.xlsx"
dob.file <- "cdr.xlsx"
```

APOE genotyping was revised in January 2019, thus APOE data extracted from an updated file supplied by Cynthia Picard:
```{r}
apoe.file <- "~/Research/Cohort_data/ADPREVENT/Patient_data/Altered_data/APOE_genotype_correction.xlsx"
```

## Libraries
```{r}
library(openxlsx)
library(tidyverse)
```

## Read & clean data

### Info (gender) data
```{r}
info.file <- paste(raw.dir, info.file, sep = "")
read.xlsx(info.file) %>%
    colnames()

info <- read.xlsx(info.file) %>%
    mutate(iid = as.factor(PSCID), sex = factor(Gender, levels = c("Male", "Female"))) %>%
    select(iid, sex) %>%
    unique() 
info.n <- length(unique(info$iid))

summary(info)
```

### APOE genotype data
```{r}
colnames(read.xlsx(apoe.file, sheet = "APOE_all"))

apoe.data <- read.xlsx(apoe.file, sheet = "APOE_all") %>%
    select(iid = Patient, apoe.geno = APOE_geno, apoe4.status = APOE4_bin)

# clean up apoe genotype
unique(apoe.data$apoe.geno)
apoe.data[apoe.data == "3-2"] <- "2-3"
apoe.data[apoe.data == "4-3"] <- "3-4"
apoe.data[apoe.data == "4-2"] <- "2-4"

apoe.data$apoe.geno <- factor(apoe.data$apoe.geno)
head(apoe.data)
```

### Date of birth data
```{r}
colnames(read.xlsx(paste(raw.dir, dob.file, sep = "")))

dob <- read.xlsx(paste(raw.dir, dob.file, sep = "")) %>%
    select(iid = PSCID, dob = "10_birthdate_date")
```

### Merge data
```{r}
m <- merge.data.frame(info, apoe.data, all = TRUE) %>%
    merge.data.frame(dob, all = TRUE) %>%
    unique()
summary(m)
```

### Check unique iids
```{r}
length(unique(m$iid)) == nrow(m)
```

## Write data
```{r}
write.table(m, file = "prad_demog_data.txt", 
            quote = FALSE, row.names = FALSE, sep = "\t")
```

