---
title: "Log of 'prad_pca_data.txt'" 
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

### Raw data source
```{r, echo=FALSE}
raw.dir <- "~/Research/Cohort_data/ADPREVENT/Patient_data/DATA_RELEASE_5.0_20171130/raw_data/dataDump01Dec17/nn_xlsx/" 
```
`r raw.dir`

Note that data are  collected from PREVENT-AD data release 5.0.

### Raw data files
Paths relative to raw data directory (files and names are exactly the same as raw data files, just converted to '.xlsx' format):
```{r}
dr.5.file <- "candidate_info.xlsx"
```

### Genetic data
```{r}
pca.file <- "~/Research/Cohort_data/ADPREVENT/Genetic_data/PRAD_2019/pca-1kG_v3_ld/PCA_EUR/PCA/PRAD_2019_EUR_PCA.eigenvec"

fam.file <- "~/Research/Cohort_data/ADPREVENT/Genetic_data/PRAD_2019/imputed/prad_2019_imp_eur.fam"
```

## Libraries
```{r}
library(tidyverse)
library(openxlsx)
```

## Data

### Data release 5.0 ids
```{r}
dr.5.ids <- read.xlsx(dr.5.file) %>%
    rename(IID = PSCID) %>%
    unique()
nrow(dr.5.ids)
summary(dr.5.ids)
unique(dr.5.ids$SubprojectID)
```
Data release 5.0 contains a "Young" cohort. These are significantly younger than the rest of the cohort and will be excluded from the study. 

```{r}
dr5.old <- filter(dr.5.ids, SubprojectID != "Young") %>%
    unique()
```
Thus, there are **`r nrow(dr5.old)`** individuals in data release 5.0 qualifying.

### PC data
```{r}
pca <- read.table(pca.file, header = TRUE)
summary(pca)
```
There are **`r nrow(pca)`** individuals with pca data.

### .fam data
```{r}
fam <- read.table(fam.file) %>%
    mutate(sex = V5 - 1) %>%
    select(IID = V2, sex)
summary(fam)
```

#### Merge data
```{r}
m <- merge.data.frame(pca, fam) %>%
    merge.data.frame(score)

m.f <- filter(m, IID %in% dr5.old$IID)
```
There are **`r nrow(m)`** individuals with both score and pca data (the one missing is due to that scoring was done with imputed data (from CP) and PC:s were calculated with QC:ed data set by NN). Of these **`r nrow(m.f)`** individuals have data in data release 5.0 (excluding the "Young" cohort).

## Standardized scores
```{r}
m.f$tc.pgs <- scale(m.f$SCORE)
```

## Write data
```{r}
write.table(m.f, file = "prad_genetic_data.txt", 
            row.names = FALSE, quote = FALSE, sep = "\t")
```

