---
title: "Log of 'prad_csf_data.txt'" 
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

### Raw data source
```{r, echo=FALSE}
raw.dir <- "~/Research/Cohort_data/ADPREVENT/Patient_data/DATA_RELEASE_5.0_20171130/raw_data/dataDump01Dec17/nn_xlsx/" 
```
`r raw.dir`

Note that data are  collected from PREVENT-AD data release 5.0.

### Raw data files
Paths relative to raw data directory (files and names are exactly the same as raw data files, just converted to '.xlsx' format):
```{r}
csf.file <- "LP.xlsx"
```

## Libraries
```{r}
library(openxlsx)
library(tidyverse)
```

## Read & clean data
```{r}
colnames(read.xlsx(paste(raw.dir, csf.file, sep = "")))

csf <- read.xlsx(paste(raw.dir, csf.file, sep = "")) %>%
    mutate(csf.age = Candidate_Age/12) %>%
    select(iid = PSCID,
           visit = Visit_label, 
           csf.age, 
           csf.date = Date_taken, 
           csf.ab42 = ELISA_final_b_amyloid1_42,
           csf.ab40 = Abeta40,
           csf.tau = ELISA_tau, 
           csf.ptau = ELISA_ptau)
knitr::kable(head(csf))
```

**Number of measurements per visit:**
```{r}
csf.g <- group_by(csf, visit) %>%
  summarise(n.ab42 = sum(!is.na(csf.ab42)),
            n.ab40 = sum(!is.na(csf.ab40)),
            n.tau = sum(!is.na(csf.tau)),
            n.ptau = sum(!is.na(csf.ptau)))
knitr::kable(csf.g)
```

## Write data
```{r}
write.table(csf, file = "prad_csf_data.txt", 
            quote = FALSE, row.names = FALSE, sep = "\t")
```

