---
title: "Log of 'prad_statin_data.txt'" 
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

### Raw data source
```{r, echo=FALSE}
raw.dir <- "~/Research/Cohort_data/ADPREVENT/Patient_data/DATA_RELEASE_5.0_20171130/raw_data/dataDump01Dec17/nn_xlsx/" 
```
`r raw.dir`

Note that data are  collected from PREVENT-AD data release 5.0.

### Raw data files
Paths relative to raw data directory (files and names are exactly the same as raw data files, just converted to '.xlsx' format):
```{r}
statin.file.bl <- "current_med_use.xlsx"
statin.file.fu <- "follow_up.xlsx"
```

## Libraries
```{r}
library(tidyverse)
library(openxlsx)
```

## Read & clean data

**Raw data columns**
```{r}
statin.file.bl <- paste(raw.dir, statin.file.bl, sep = "")
read.xlsx(statin.file.bl) %>%
  colnames()
```

Baseline data:
```{r}
statin.bl <- read.xlsx(statin.file.bl) %>%
  select(iid = PSCID, 
         visit = Visit_label,
         drug.date = Date_taken, 
         contains("reductase")) %>%
    gather("col.head", "comment", 4:9)
```

Follow up data:
```{r}
statin.file.fu <- paste(raw.dir, statin.file.fu, sep = "")
statin.fu <- read.xlsx(statin.file.fu) %>%
  select(iid = PSCID, 
         visit = Visit_label,
         drug.date = Date_taken, 
         contains("reductase")) %>%
    gather("col.head", "comment", 4:12)
```

Merge data:
```{r}
m <- rbind.data.frame(statin.bl, statin.fu)
unique(m$col.head)

m$col.head <- gsub("5_", "", m$col.head)
m$col.head <- gsub("2_e_", "", m$col.head)

unique(m$col.head)

# Types of statins used:
unique(m$comment)

# Create statin column
m$statin <- lapply(m$comment, function(x)
    if(is.na(x) | x == "not_answered") {NA}
    else if(x == "none") {0}
    else {1}) %>%
    unlist()


discontinued <- filter(m, grepl("_discontinued", col.head)) %>%
    filter(statin == 1)

sustained <- filter(m, !grepl("_discontinued", col.head) & !is.na(statin)) %>%
    select(iid, visit, statin) %>%
    unique() %>%
    group_by(iid, visit) %>%
    summarize(sum.statin = sum(statin)) %>%
    spread(visit, sum.statin)
summary(sustained)
```

Statin data are mostly present for the "PREEL00" visit, the other visits are missing too many data points to be useful. Thus, selecting statin data only for "PREEL00" visit.

```{r}
statin.final <- filter(m, visit == "PREEL00") %>%
    select(iid, visit, drug.date) %>%
    merge.data.frame(sustained) %>%
    select(iid, visit, statin.date = drug.date, statin = PREEL00) %>%
    unique()
```

## Write data
```{r}
write.table(statin.final, file = "prad_statin_data.txt", 
            row.names = FALSE, quote = FALSE, sep = "\t")
```

