---
title: "Log of 'prad_lipid_data.txt'" 
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

### Raw data source
```{r, echo=FALSE}
raw.dir <- "~/Research/Cohort_data/ADPREVENT/Patient_data/DATA_RELEASE_5.0_20171130/raw_data/dataDump01Dec17/nn_xlsx/" 
```
`r raw.dir`

Note that data are  collected from PREVENT-AD data release 5.0.

### Raw data files
Paths relative to raw data directory (files and names are exactly the same as raw data files, just converted to '.xlsx' format):
```{r}
data.file <- "lab.xlsx"
```

## Libraries
```{r}
library(openxlsx)
library(tidyverse)
```

## Read & clean data

**Raw data columns**
```{r}
data.file <- paste(raw.dir, data.file, sep = "")
read.xlsx(data.file) %>%
  colnames()
```

**Filtered data**
```{r}
tc.data <- read.xlsx(data.file) %>%
    mutate(tc.age = Candidate_Age/12) %>%
  select(iid = PSCID, 
         visit = Visit_label, 
         tc.age, 
         tc.date = Date_taken, 
         tg.mM = lipid_triglycerides_result, 
         tc.mM = lipid_cholesterol_result, 
         ldl.mM = lipid_LDL_result, 
         hdl.mM = lipid_HDL_result)
```

Non-HDL cholesterol levels calculated as TC - HDL according to Langsted & Nordestgaard 2019 (Pathology)
```{r}
tc.data$nonhdl.mM = tc.data$tc.mM - tc.data$hdl.mM

head(tc.data)

summary(tc.data)
```

**Number of measurements per visit**
```{r}
tc.data.g <- group_by(tc.data, visit) %>%
  summarise(n.tg = sum(!is.na(tg.mM)), 
            n.tc = sum(!is.na(tc.mM)),
            n.ldl = sum(!is.na(ldl.mM)),
            n.hdl = sum(!is.na(hdl.mM)), 
            n.nonhdl = sum(!is.na(nonhdl.mM)))
tc.data.g
```

## Write data
```{r}
write.table(tc.data, file = "prad_lipid_data.txt", 
            quote = FALSE, row.names = FALSE, sep = "\t")
```

