---
title: "Log of 'prad_age_data.txt'" 
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

### Raw data source
```{r, echo=FALSE}
raw.dir <- "~/Research/Cohort_data/ADPREVENT/Patient_data/DATA_RELEASE_5.0_20171130/raw_data/dataDump01Dec17/nn_xlsx/" 
```
`r raw.dir`

Note that data are  collected from PREVENT-AD data release 5.0.

### Raw data files
Paths relative to raw data directory (files and names are exactly the same as raw data files, just converted to '.xlsx' format):
```{r}
bl.file <- "baseline.xlsx"
fu.file <- "follow_up.xlsx"
dob.file <- "cdr.xlsx"
mri.file <- "mri_feedbacks_ASL.xlsx"
```

## Libraries
```{r}
library(openxlsx)
library(tidyverse)
```

## Read & clean data

### Baseline data
```{r}
colnames(read.xlsx(paste(raw.dir, bl.file, sep = "")))[1:20]
```

```{r}
bl <- read.xlsx(paste(raw.dir, bl.file, sep = "")) %>%
    select(iid = PSCID, visit = Visit_label, Candidate_Age, date = Date_taken) %>%
    mutate(age = Candidate_Age/12) %>%
    unique()
summary(bl)
```

### Follow up data
```{r}
colnames(read.xlsx(paste(raw.dir, fu.file, sep = "")))[1:20]
```

```{r}
fu <- read.xlsx(paste(raw.dir, fu.file, sep = "")) %>%
    select(iid = PSCID, visit = Visit_label, Candidate_Age, date = Date_taken) %>%
    mutate(age = Candidate_Age/12)
summary(fu)
```

### DOB data
```{r}
colnames(read.xlsx(paste(raw.dir, dob.file, sep = "")))

dob <- read.xlsx(paste(raw.dir, dob.file, sep = "")) %>%
    mutate(dob = as.Date(`4_birthdate_date`)) %>%
    select(iid = PSCID, dob) %>%
    unique()
```

### MRI file
```{r}
colnames(read.xlsx(paste(raw.dir, mri.file, sep = "")))

mri <- read.xlsx(paste(raw.dir, mri.file, sep = "")) %>%
    mutate(date = as.Date(MRI_Acquisition_Date)) %>%
    select(iid = PSCID, visit = Visit_label, date) %>%
    merge.data.frame(dob) %>%
    mutate(age = as.numeric(as.character((date - dob)/365.25)))
```

### Combine data
```{r}
m <- rbind.data.frame(bl, fu) %>%
    unique() %>%
    select(-Candidate_Age) %>%
    merge.data.frame(mri, by = c("iid", "visit"), suffixes = c(".v", ".mri"), all = TRUE) 

# calculate new age
m$age <- apply(select(m, age.v, age.mri), 1, function(x) if(!is.na(x[1])) {x[1]} else {x[2]})

m.l <- select(m, iid, visit, age) %>% unique()

m.w <- spread(m.l, key = visit, value = age)
sum(duplicated(m.w$iid))
```

## Write data
```{r}
write.table(m.l, file = "prad_age_data_long.txt", 
            quote = FALSE, row.names = FALSE, sep = "\t")
write.table(m.w, file = "prad_age_data_wide.txt", 
            quote = FALSE, row.names = FALSE, sep = "\t")
```

