---
title: "Log of 'prad_bl_data.txt'" 
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

### Raw data source
```{r, echo=FALSE}
raw.dir <- "~/Research/Cohort_data/ADPREVENT/Patient_data/DATA_RELEASE_5.0_20171130/raw_data/dataDump01Dec17/nn_xlsx/" 
```
`r raw.dir`

Note that data are  collected from PREVENT-AD data release 5.0.

### Raw data files
Paths relative to raw data directory (files and names are exactly the same as raw data files, just converted to '.xlsx' format):
```{r}
bl.file <- "baseline.xlsx"
```

## Libraries
```{r}
library(openxlsx)
library(tidyverse)

source(paste(path.to.project, "scripts/project_functions.R", sep = ""))
```

## Read & clean data

### Info (gender) data
```{r}
read.xlsx(paste(raw.dir, bl.file, sep = "")) %>%
    colnames()

bl <- read.xlsx(paste(raw.dir, bl.file, sep = "")) %>%
    mutate(iid = as.factor(PSCID), visit = Visit_label, bl.date = Date_taken, bl.age = Candidate_Age) %>%
    select(iid, visit, bl.date, bl.age) %>%
    unique() 
bl.n <- length(unique(bl$iid))

summary(bl)
```

Some individuals have multiple baseline values:
```{r}
filter(bl, iid %in% bl$iid[duplicated(bl$iid)]) %>%
    arrange(iid)
```
```{r}
no.dup <- filter(bl, !iid %in% bl$iid[duplicated(bl$iid)])
dup <- filter(bl, iid %in% bl$iid[duplicated(bl$iid)]) %>%
    filter(visit == "PREBL00")
bl.final <- rbind.data.frame(no.dup, dup)
summary(bl.final)
```

## Write data
```{r}
write.table(bl.final, file = "prad_bl_data.txt", 
            quote = FALSE, row.names = FALSE, sep = "\t")
```

