---
title: "Log of 'prad_pet_data.txt'" 
subtitle: "Project: thesis_tc_pgs_in_AD"
author: "Nathalie Nilsson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    df_print: kable
    fig_height: 5.2
    fig_width: 6.1
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.wd <- gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```
**Location of project directory:** `r path.to.project`

**Working directory relative to project:** `r path.to.wd`

## Input

### Raw data source
PET data were supplied by Sylvia Villeneuve's lab. 

### Raw data files
```{r}
pet.file <- "~/Research/Cohort_data/ADPREVENT/Patient_data/from_sv_lab/Amyloid_tau_PET_Nathalie.xlsx"
```

## Libraries
```{r}
library(tidyverse)
library(openxlsx)
```

## Read & clean data
```{r}
colnames(read.xlsx(pet.file, rows = c(1, 3:396), detectDates = TRUE))

pet <- read.xlsx(pet.file, rows = c(1, 3:396), detectDates = TRUE) %>%
    select(iid = PSCID, 
           visit = Label_PET_visit, 
           petab.date = Date_AmyloidPET_visit, 
           petab = Global_Amyloid_n129, 
           pettau.date = Date_TauPET_visit,
           pettau.etx = tau_ctx_mean_entorhinal_n129, 
           pettau.meta = tau_metaROI_RikOssenkoppele, 
           pettau.pos = Tau_positive_entorh_VilleneuveLab_n129)
summary(pet)

unique(pet$visit) 
```

Filter data to only contain IDs with either petab or pettau data
```{r}
pet$pet.data <- apply(pet[, c("petab", "pettau.etx")], 1, function(x) {
    if(!is.na(x[1]) | !is.na(x[2])) {1}
    else {0}
})

pet.f <- filter(pet, pet.data == 1)
```

**Check duplicated ids**
```{r}
sum(duplicated(pet.f$iid))
```

There is **`r nrow(pet.f)`** individuals with at least one PET visit (these do not correspond to visit labels baseline).

## Write data
```{r}
write.table(pet.f, file = "prad_pet_data.txt", 
            quote = FALSE, row.names = FALSE, sep = "\t")
```

